R Notebook

### 2020-01-23 src/comb-maxent.Rmd
fuck you, fuck you all fuckers

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(prback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  #mess_pr <- subset(train, occurrence == 1)
  #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  if (yr == "1998"){
    betaval <- 0.25
  } else if (yr == "2001"){
    betaval <- 0.75
  } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
    betaval <- 1
  } else if (yr == "2005"){
    betaval <- 1.25
  } else if (yr == "1999" | yr == "2000"| yr == "2009"){
    betaval <- 1.5
  } else if (yr == "2007"){
    betaval <- 1.75
  } else if (yr == "2010"){
    betaval <- 2
  } else if (yr == "2008"){
    betaval <- 2.25
  }

    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

    saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))

    #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

    #test auc scores and other lovely things you might want to include
    ev <- evaluate(p = testp, a= testb, model = mod)
    evauc <- ev@auc
    evcor <- ev@cor
    evpcor <- ev@pcor
    evnp <- ev@np
    evna <- ev@na
    evmaxtprtnr <- threshold(ev, 'spec_sens')
    evmaxkap <- threshold(ev, 'kappa')
    evno_omission <- threshold(ev, 'no_omission')
    evprevalence <- threshold(ev, 'prevalence')
    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
    evsensitivity <- threshold(ev, 'sensitivity')

    #continuous boyce index
    predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
    predPres <- predPres$Cloglog.prediction
    predBg <- predBg$Cloglog
    source("../src/contBoyce.R")
    cbi <- contBoyce(pres = predPres, bg = predBg)

      aicetc$train_AUC <- trainauc #training only AUC - not tested
      aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
      aicetc$coefficients <- nocoeff #this is to calculate the AIC
      aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
      aicetc$test_AUC <- evauc #testing AUC
      aicetc$test_evcor <- evcor #testing correlation coefficient
      aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
      aicetc$test_nopresence <- evnp 
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$tss <- max(ev@TPR + ev@TNR) - 1
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval
      aicetc$cbi <- cbi

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))

    #calculate minimal predicted area using ecospat
    aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
    aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
    aicetc$year <- yr

    write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

@@

clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/dec"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[6])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc2, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

@@

write.csv(cumst1, "../output/maxent/paper_plots/prob_by_depth/cellsperlayermar.csv", row.names = FALSE)

>>>>

# 
# 
# for (year in 1:nrow(modelmonth)){
#   timeslice <- modelmonth[year] #testing timeslice
#   yr <- str_sub(timeslice, -4) #the year being used for testing
#   train <- subset(prback, year != yr) #training data is all years but testing
#   trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
#   test <-  subset(prback, year == yr) #creates the testing dataset
#   testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
#   testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
#   testb <- subset(test, occurrence == "0") #splits the testing dataset into background
#   testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
#   #mess_pr <- subset(train, occurrence == 1)
#   #mess_pr  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
#   train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling
# 
#   
#   if (yr == "1998"){
#     betaval <- 0.25
#   } else if (yr == "2001"){
#     betaval <- 0.75
#   } else if (yr == "2002" | yr == "2003"| yr == "2004"| yr == "2006"| yr == "2011"){
#     betaval <- 1
#   } else if (yr == "2005"){
#     betaval <- 1.25
#   } else if (yr == "1999" | yr == "2000"| yr == "2009"){
#     betaval <- 1.5
#   } else if (yr == "2007"){
#     betaval <- 1.75
#   } else if (yr == "2010"){
#     betaval <- 2
#   } else if (yr == "2008"){
#     betaval <- 2.25
#   }
# 
# 
#   
#   
#     mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/models/test-comb/", timeslice), args=c("writebackgroundpredictions=TRUE",  "maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))
#     
#     saveRDS(mod, file= paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
# 
# 
#     #now based on glover-kapfer pdf with excel manual + https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/10-1171.1... + adaptation for        getting testAUC.. horrible code
# 
#     #count number of non-zero coefficients, excluding the final four rows
#     coefile <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
#     coefile <- head(coefile, -4)
#     coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
#     nocoeff <- nrow(coefile2)
#     bick <- sum(as.numeric(coefile2[, 2]))
#     bic <- (bick*log(sum(trainpa == 1)) - 2*11)
#     
#     #no presences used in model (less those used for testing)
#     maxresults <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/maxentResults.csv"), header = TRUE)
#     notrainingpoints <- maxresults$X.Training.samples
#     testauc <- maxresults$Test.AUC
#     trainauc <- maxresults$Training.AUC
#     
#     #natural log
#     samplepred <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
#     samplepred$nat_log <- log(samplepred$Raw.prediction)
#     sumsamplepred <- sum(samplepred$nat_log)
#   
#   
#     #test auc scores and other lovely things you might want to include
#     ev <- evaluate(p = testp, a= testb, model = mod)
#     evauc <- ev@auc
#     evcor <- ev@cor
#     evpcor <- ev@pcor
#     evnp <- ev@np
#     evna <- ev@na
#     evmaxtprtnr <- threshold(ev, 'spec_sens')
#     evmaxkap <- threshold(ev, 'kappa')
#     evno_omission <- threshold(ev, 'no_omission')
#     evprevalence <- threshold(ev, 'prevalence')
#     evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
#     evsensitivity <- threshold(ev, 'sensitivity')
#     
#     #continuous boyce index
#     predPres <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
#     predBg <- read.csv(paste0("../output/maxent/models/test-comb/", timeslice, "/species_backgroundPredictions.csv"), header = TRUE)
#     predPres <- predPres$Cloglog.prediction
#     predBg <- predBg$Cloglog
#     source("../src/contBoyce.R")
#     cbi <- contBoyce(pres = predPres, bg = predBg)
#     
#   
#       aicetc$train_AUC <- trainauc #training only AUC - not tested
#       aicetc$no_trainng_points <- notrainingpoints #numberof points used for training
#       aicetc$coefficients <- nocoeff #this is to calculate the AIC
#       aicetc$sum.of.logs <- sumsamplepred #this is to calculate the AIC
#       aicetc$test_AUC <- evauc #testing AUC
#       aicetc$test_evcor <- evcor #testing correlation coefficient
#       aicetc$test_evpcor <- evpcor # testing p-value for correlation coefficient
#       aicetc$test_nopresence <- evnp 
#       aicetc$test_nobackground <- evna
#       aicetc$test_maxsumsenspec <- evmaxtprtnr
#       aicetc$test_maxkap <- evmaxkap
#       aicetc$test_no_omission <- evno_omission
#       aicetc$test_prevalence <- evprevalence
#       aicetc$test_equal_sens_spec <- evequal_sens_spec
#       aicetc$test_sensitivity <- evsensitivity
#       aicetc$tss <- max(ev@TPR + ev@TNR) - 1
#       aicetc$test_bic <- bic #not sure this is correct...
#       aicetc$beta <- betaval
#       aicetc$cbi <- cbi
#       
#     aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
#     aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$no_trainng_points/(aicetc$no_trainng_points - aicetc$coefficients-1))))
#     
#     #calculate minimal predicted area using ecospat
#     aicetc$mpa_0.9_cum <- ecospat.mpa(samplepred$Cumulative.prediction, perc = 0.9) #90%
#     aicetc$mpa_0.9_clog <- ecospat.mpa(samplepred$Cloglog.prediction, perc = 0.9) #90%
#     aicetc$year <- yr
# 
# 
#     write.csv(aicetc, paste0("../output/maxent/models/test-comb/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

@@

clogasc_list <- list.files(paste0("../output/maxent/predictions/com-split/mar"), pattern = "^avg_cloglog_.*.asc", full.names = TRUE) #just choose a random month to calculate cells
cumst <- data.frame(0, 0)
colnames(cumst) <- c("depth", "no_cells")
for (cum in 1:length(clogasc_list)){
  cumasc <- raster(clogasc_list[cum])
  ascname <- names(cumasc)
  cumst[cum, 1] <- str_sub(ascname, 13)
  cumasc2 <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
  ccnt <- freq(cumasc2, useNA='no')
  ccnt <- sum(ccnt[ ,2])
  cumst[cum, 2] <- ccnt
}

@@

write.csv(cumst1, "../output/maxent/paper_plots/cellsperlayer.csv", row.names = FALSE)

### 2020-01-23 src/testing_file.Rmd
fuck you, fuck you all fuckers

>>>>

#need to reclassify cells outside the mcp to NA for predictions
```{r}
outsidemcp <- shapefile("../data/env/landp/outside-mcp-wgs84.shp") #made in arcgis - in aea

ascdirs <- list.dirs("../data/env/ascii-original", full.names = FALSE, recursive = FALSE)

for (ad in 1:length(ascdirs)){
  timeslice <- ascdirs[ad]
  ascfiles <- list.files(paste0("../data/env/ascii-original/", timeslice), pattern = "\\.asc$", full.names = TRUE)
  for (file in 1:length(ascfiles)){
    ras<- raster(ascfiles[file])
    crs(ras) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "
    ascname <- str_sub(ascfiles[file], start = 38)
    ras[outsidemcp] <- 500000
    ras[ras == 500000] <- NA
    writeRaster(ras, paste0("../data/env/asciitrim/", ascdirs[ad], "/", ascname), overwrite=TRUE)
  }
}

test <- raster("../output/maxent/predictions/com-split/apr/cloglog_1998_04_0.50576.asc") #just chosen randomly to act as dummy file
crs(test) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0 "

test[outsidemcp] <- 500000
test[test == 500000] <- NA
plot(test)

#if need to predict from the model...

#monthlist <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec") # a list of months to order by

modlist <- list.dirs("../output/maxent/models/test-comb", full.names = FALSE, recursive = FALSE) 
#modlist <- modlist[ grepl(mthna, modlist) ]

for (mth in 1:length(modlist)){
    mthna <- str_sub(modlist[mth], end = 3)
    yr <- str_sub(modlist[mth], start = 4)
    timeslice <- modlist[mth]
    mod <- readRDS(paste0("../output/maxent/models/test-comb/", timeslice, "/", timeslice, ".rda"))
    unlayer <- read.csv(paste0("../data/env/test-comb/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

    unlayer_yr <- subset(unlayer, ascyearslist == yr)
    #unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname))
    stackdeplst[[yr]] <- stackdep
    write.csv(stackdep, paste0("../output/maxent/models/test-comb/", timeslice, "/stackdepthlist_", yr, ".csv"), row.names = FALSE)

    for (td in 1:nrow(unlayer_yr)){
        temp_depth <- raster(unlayer_yr$temp_depth[td])
        crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(temp_depth) <- "temp_depth"
        #temp_depth <- projectRaster(temp_depth, aea)
        chl_surface <- raster(unlayer_yr$chl_surface[td])
        crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(chl_surface) <- "chl_surface"
        #chl_depth <- projectRaster(chl_depth, aea)
        o2_depth <- raster(unlayer_yr$o2_depth[td])
        crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(o2_depth) <- "o2_depth"
        #o2_depth <- projectRaster(o2_depth, aea)
        salinity_depth <- raster(unlayer_yr$salinity_depth[td])
        crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(salinity_depth) <- "salinity_depth"
        #salinity_depth <- projectRaster(salinity_depth, aea)
        nao_sample <- raster(unlayer_yr$nao_sample[td])
        crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(nao_sample) <- "nao_sample"
        #nao_sample <- projectRaster(nao_sample, aea)
        nao_prev <- raster(unlayer_yr$nao_prev[td])
        crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(nao_prev) <- "nao_prev"
        #nao_prev <- projectRaster(nao_prev, aea)
        nao_winter <- raster(unlayer_yr$nao_winter[td])
        crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
        names(nao_winter) <- "nao_winter"
        #nao_winter <- projectRaster(nao_winter, aea)
        amo_sample <- raster(unlayer_yr$amo_sample[td])
        crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
        names(amo_sample) <- "amo_sample"
        #amo_sample <- projectRaster(amo_sample, aea)
        amo_winter <- raster(unlayer_yr$amo_winter[td])
        crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
        names(amo_winter) <- "amo_winter"
        #amo_winter <- projectRaster(amo_winter, aea)
        stackn <- unlayer_yr$stackname[td]
        stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)

        #stackSave(test,unlayer_yr$stackname[td])
        pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=cloglog')) 
        # predraw <- predict(mod, stack, file = paste0("../output/maxent/predictions/test-comb/", mthna, "/", "raw_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=raw')) 
        prplot <- plot(pred, png(paste0("../output/maxent/predictions/test-comb/", mthna, "/", "cloglog_", stackn, ".png")))
        dev.off() # stops automatic saving of the plot to a png
    }
}
```
